%!TEX root = ./Grimdark.tex
\chapter{Character Creation}
Grimdark characters are created by establishing the characters Life Path. The process will guide you through a series of steps which mirror certain phases of a characters (early) life. You will pick up Attribute modifications, Aptitudes, Professions and skills along this path, which will define your starting character in addition to giving everyone a good impression about the background of the character.\\
Below you will find a table summarizing the life path steps and what you will get from each.

\begin{DndTable}[header=Life Path Overview]{llcl}
  \textbf{Step}  & \textbf{Attributes} & \textbf{Skills} & \textbf{Other}\\
	\nameref{Birthright} & Baseline & - & Limits \\
	\nameref{sec:lifepath_2_Homeworld} & +3/+3/-3 & 3 & Aptitude 1\\
	\nameref{sec:lifepath_3_childhood} & +3 & 2 & \parbox[t]{2.5cm}{1 Profession,\\Aptitude 2} \\
	\nameref{sec:lifepath_4_education} & +3/-3 & 4 & \parbox[t]{2.5cm}{2 Professions,\\Group Aptitude}\\
	\nameref{sec:lifepath_5_combat} & - & 3 & -\\
	\nameref{sec:lifepath_6_early_career} & +3/+3 & 3 & 1 Profession\\
\end{DndTable}

\subsection{Creation Rules \& Constraints}
	Due to the vast amount of possible combinations allowed by the life path system, here are some basic rules and constraints:
	\begin{itemize}
		\item \textbf{Free Skill Points:} In some stages, you may gain one or many Free Skill Points. These can be used to raise any Skill you want, which isn't restricted and marked with a * (see \ref{skill_table}).
		\item \textbf{Any X:} This is a shorthand for a long list of options open for you. X may refer to different things depending on the context. Most commonly it will refer to a Profession in the context of Skill selection, which means you can pick any one Skill associated with that Profession.
		\item \textbf{Profession and Skills:} Whenever you gain a Profession, you will also gain a number of skill points to spend within the Profession. You must spend these points on Skills associated with the gained Profession.
		\item \textbf{Duplicate Aptitude:} You can not get the same Aptitude twice. If the step that would give you a duplicate provides a choice, pick another option. Otherwise pick any other Aptitude (not a Group Aptitude).
		\item \textbf{Duplicate Skill:} Increase the skill level to the next higher. No skill can be raised beyond the second level ("Experienced", see: \href{skill_level}). Pick another level if you would bump a skill to the third level.
		\item \textbf{Duplicate Profession:} Each Profession you gain during character creation comes with a number of skill points to spent within the Profession. Skill points from duplicated Professions are reduced by one but can otherwise be freely spent in the respective Profession.
		\item \textbf{Attribute Scores:} Attribute scores may not be raised beyond a characters limit. Note that due to the basic resolution mechanic used in Grimdark, raising a single Attribute score while neglecting all others is not a good idea. You should at least try to strike a balance between the two most important Attributes of your character.

	\end{itemize}

\section{Step 1: Birthright} % (fold)
	\label{Birthright}
	Some things are determined at birth. Are you a baseline human or an abhuman? Or even a Xeno or mutant - \textit{someone call the Inquisition!}
	Your birthright depends your baseline Attributes and also your natural limits, that is the highest your attributes can get with training and experience. Raising them above your limits, will require special unnatural aids, like technology, drugs or even mutations and warp influence.

	The baseline Attributes are determined by rolling a number of five-sided dice and adding a fixed value on top of the result. This is denoted by xD5+y, where x determines the number of dice rolled and y denotes the fixed value counted on top. X and Y are determined by your birthright. Whether you roll the entire set and assign the results to the Attributes or roll them in order is a matter of preference or your GM's call.
	The first has a more natural feel to it, while the latter works better if you already have a certain character concept in mind at the start of creation.

	Make sure to check in with your GM which Birthrights are available in your campaign before starting - he may after all be unwilling to include abhumans in a game of noble intrigue or wants to run with everyone playing an Ogryn!

	\begin{DndTable}[header=Birthright Attributes]{XXX}
	    \textbf{Attribute}  & \textbf{Baseline Human} & \textbf{Ogryn}\\
		Power & 4D5+20 & 4D5+40\\
		Finesse & 4D5+20 & 2D5+10\\
		Toughness &  4D5+20 & 4D5+40\\
		Instinct & 4D5+20 & 4D5+20\\
		Smarts & 4D5+20 & 2D5+10\\
		Will & 4D5+20 & 4D5+20\\
		Charisma & 4D5+20 & 2D5+20\\
	\end{DndTable}

	\begin{DndSidebar}{Roll-in-Order vs. Roll-and-Assign}
	You can either roll your Attributes in order of appearance or roll the entire set and then assign the results to your Attributes.
	The first method has a more natural feel to it and can lead you down character creation when you don't have a clear concept in mind at start. The second variant works better if you have a certain theme or concept in mind for your character. If you want to play a hulking gladiator pit fighter and you roll low for his Power Attribute, your prepared concept might not come out as imagined.

	Grimdark does not forces one or the other (but your GM may of course). Choose whatever works best for you.
	As an optional method, groups looking for a more balance Attribute generation may opt to take the average values for the baseline. To do so, simply count each D5 you would normally roll as a three. So 4D5+20 (the baseline human roll) will give a result of 32 ($4*3=12 + 20 = 32$).
	\end{DndSidebar}

	\subsection{Baseline Human}
	The most common Birthright in the Imperium by far. The vast majority of Imperial citizens will never see an abhuman or mutant in their life's - God Emperor bless their ignorance.\\
	Their baseline Attributes all are determined by rolling four five sided dice (4D5) and add 20 to the result.
	Their Limits are at 50 for Power and Will and 60 for all other Attributes.
	% section section_name (end)

\section{Step 2: Homeworld} % (fold)
	\label{sec:lifepath_2_Homeworld}
	The place of ones birth as a lasting impact on a character - although in a wide variety of ways. After all, a Hive world will leave a very different impact in a highborn, compared to an underhiver.\\
	In this stage, you will get your first Attribute adjustments (+3 to two, -3 to a third), your first Aptitude and three skills.
	Available Homeworlds are:
	\begin{itemize}
		\item Deathworld (Primitive)
		\item Deathworld (Civilized)
		\item Backwater
		\item Civilized World
		\item Forge World
		\item Hive World
		\item Fortress World
		\item Agriworld
		\item Shrine World
		\item Voidborn
	\end{itemize}


	\subsection{Primitive Deathworld}
	You were born on a primitive deathworld. No technology to Ward the extreme environment. No weaponry more advanced than the crudest of blackpowder weapons to fend off the local wildlife (and population). Anyone surviving this, will be a hardy fellow, able to make the most out of what little he or she is given.

	\begin{itemize}
		\item Aptitude (pick one): Power, Instinct, Toughness
		\item Attributes: 
		\begin{itemize}
			\item Toughness+3
			\item Instinct+3 or Will+3
			\item Smarts-3 or Finesse-3
		\end{itemize}
		\item Skills (pick 3): any Observer, any Ghost, any Believer, any Physical (excluding Gunman skills)
	\end{itemize}

	\subsection{Civilized Deathworld}
	You were born on a deathworld, which has the common imperial tech level at its disposal - and yet is still classified a deathworld. Maybe the environment is hideously toxic and life is only possible within shielded domes of steel and energy. Maybe the flora and fauna is putting up such a commendable fight, that colonization of the world has never truly succeeded beyond some hot-spots of human occupation. Life is hard but at least you (probably) had a home with the most basic installations for warmth, water and electricity. Most place like this also tend to instill a rather close-knit bond between its citizens. 

	\begin{itemize}
		\item Aptitude (pick one): Toughness, Charisma, Instincts
		\item Attributes:
		\begin{itemize}
			\item Charisma+3
			\item Instinct+3 or Toughness+3
			\item Power-3 or Finesse-3
		\end{itemize}
		\item Skills: 
		\begin{itemize}
		 	\item Pick two: Imperium of Man (Loremaster), any Operator, any Technomat
		 	\item One free skill point
		 \end{itemize} 
	\end{itemize}

	\subsection{Backwater}
	Being born on a backwater world means facing disregard and disinterest on a planetary-system-wide level. The wider Imperium doesn't really care for the world, let alone its inhabitants. Maybe the world does not have any natural resources (left) or it is simply too far off any (safe) trading routes to attract merchants. Backwater worlds thus usually fall into one of two extremes: either they are places of heightened religious worship trying to make up their lack of physical wealth with spiritual one or they are crime-ridden, lawless places ruled by shadowy figures using local politicians as their marionettes. 

	\begin{itemize}
		\item Aptitude (pick one): Finesse, Charisma, Will
		\item Attributes:
		\begin{itemize}
			\item Charisma+3
			\item Will+3 or Finesse+3
			\item Smarts-3
		\end{itemize}
		\item Skills: 
		\begin{itemize}
		 	\item Imperium of Man (Loremaster) or Shadow Organizations (Loremaster)
		 	\item Pick one: Technomat (Tinkerer), Bribe (Trader), Haggle (Trader), any Believer, any Betrayer, any Operator
		 	\item One free skill point
		 \end{itemize} 
	\end{itemize}


	\subsection{Civilized World}
	The most common category of imperial planets - and the most diverse. Planets of this category are mostly self-sufficient and provide a decent tithe to the greater Imperium in the form of produced goods, Imperial Guard regiments and local psykers.

	\begin{itemize}
		\item Aptitude (pick one): Smarts, Finesse, Charisma, Instincts
		\item Attributes:
		\begin{itemize}
			\item Any+3
			\item Any+3
			\item Will-3 or Toughness-3
		\end{itemize}
		\item Skills: 
		\begin{itemize}
		 	\item Imperium of Man (Loremaster) or Technomat(Tinkerer) or any Operator
		 	\item Two free skill points
		 \end{itemize} 
	\end{itemize}

	\subsection{Forge World}
	Being born on the beating, pounding heart of an Adeptus Mechanicus Forge World is being surrounded by servitors, augmented menials and Magi more machine than flesh. You were likely facing intensive schooling or actual hypno-indoctrination by learning machines and early drafting into the working force. Persons with above-average intellect and technical affinity are commonly lead to the training centers to join the ranks of the Cult Mechanicus or its auxilia. Others are worked hard, in precisely timed shifts around the clock, always teetering on the edge of their physical abilities.

	\begin{itemize}
		\item Aptitude (pick one): Smarts, Toughness
		\item Attributes:
		\begin{itemize}
			\item Smarts+3
			\item Finesse+3 or Toughness+3
			\item Will-3 or Charisma-3
		\end{itemize}
		\item Skills: 
		\begin{itemize}
		 	\item Gain the Technomat(Tinkerer) or Ways of Mars (Loremaster) skill
		 	\item Pick two: any Rationalist, any Operator, any Believer, any Loremaster
		 \end{itemize} 
	\end{itemize}

	\subsection{Hive World}
	Born amidst billion upon billion of people, hivers make up the vast majority of the Imperiums total population. Not only are they used to fill the vast ranks of soldiers and naval ratings - they also work the myriad manufactoria of their worlds in never-ending shifts to produce the goods required by the Imperium - from packed ration bars to bullets, lasguns and toilet paper. In a world where everyone can be replaced by a hundred others at any moment, the struggle for survival and keeping ones social standing is all too real and the potential depth of a downfall very deep indeed.

	\begin{itemize}
		\item Aptitude (pick one): Smarts, Toughness
		\item Attributes:
		\begin{itemize}
			\item Pick two: Finesse+3, Power+3, Smarts+3, Charisma+3
			\item Instinct-3 or Charisma-3 or Will-3 or Toughness-3
		\end{itemize}
		\item Skills: 
		\begin{itemize}
		 	\item Imperium of Man (Loremaster) or Shadow Organizations (Loremaster) or Navigate Enclosed (Navigator) or Technomat(Tinkerer)
		 	\item Selfishness(Betrayer), Endure(Athlete) or Scrutiny(Observer)
		 	\item any Operator, any Social
		 \end{itemize} 
	\end{itemize}

	\subsection{Agriworld}
	The corn baskets of the Imperium, producing unbelievable amounts of foodstuff to be exported to feed worlds unable to do themselves, e.g. Hive Worlds and Forgeworlds. Usually these worlds have a low population density in favor of wide ranges of farmland, game forests and fish-filled rivers and oceans. You probably were raised stalking game animal in dense forests, feed the endless herds of farm animal or operate massive field machines cutting down corn - long before learning to write (if you ever did).

	\begin{itemize}
		\item Aptitude (pick one): Power, Instinct
		\item Attributes:
		\begin{itemize}
			\item Instinct+3 or Charisma+3
			\item Finesse+3 or Power +3
			\item Smarts-3
		\end{itemize}
		\item Skills: 
		\begin{itemize}
		 	\item Pick three: Sixth Sense(Observer), Ghost Move(Ghost), Navigate Surface(Navigator), any Operate, any Survivalist, any Athlete, any Tinkerer, any Believer
		 \end{itemize} 
	\end{itemize}

	\subsection{Shrine World}
	Hailing from a holy shrine world, your earliest memory is likely to be some kind of prayer - by your fervent parents or played over ever present loud haulers. Wherever you go, there were signs of religious fervor: statues of saints, street musicians playing religious anthems, stained glass windows depicting the Emperor or some saint or hero and endless streams of pilgrims. Growing up in such an environment either left you deeply religious or utterly selfish.

	\begin{itemize}
		\item Aptitude (pick one): Will, Charisma
		\item Attributes:
		\begin{itemize}
			\item Will+3
			\item Charisma+3 or Power+3
			\item Smarts-3 or Finesse-3
		\end{itemize}
		\item Skills: 
		\begin{itemize}
		 	\item Any Believer or any Betrayer
		 	\item Any Trader, any Spokesman, any Enforcer
		 	\item One free skill point
		 \end{itemize} 
	\end{itemize}

	\subsection{Fortress World}
	Fortress worlds are dedicated roadblocks to contain, repel, deter or slow down grave (typically Xeno) threats from more vulnerable imperial worlds. Riddled with fortifications and supply depots, its population drilled and equipped for combat - from cradle to crypt - these worlds are hard nuts to crack for any foe and provide the wider Imperium with soldiers who know that in the grim, dark future there is only war...

	\begin{itemize}
		\item Aptitude (pick one): Will or Finesse
		\item Attributes:
		\begin{itemize}
			\item Finesse+3
			\item Power+3 or Will+3
			\item Charisma-3 or Smarts-3
		\end{itemize}
		\item Skills: 
		\begin{itemize}
		 	\item Pick two: any Warrior, any Gunman, any Operator, Tactica Imperialis (Loremaster)
		 	\item First Aid (Medic), Calculus(Rationalist), Bookkeeping(Trader), any Tinkerer, any Commander, any Physical
		 \end{itemize} 
	\end{itemize}

	\subsection{Void Born}
	You were born aboard a void ship or a deep void station and likely never have set foot on a planet for your early life years - or ever! Getting around confusing, same-looking man-made places and dealing with machines is pretty common for you.
	\begin{itemize}
		\item Aptitude (pick one): Finesse, Instinct
		\item Attributes:
		\begin{itemize}
			\item Instinct+3
			\item Finesse+3 or Power +3
			\item Toughness-3 or Charisma-3
		\end{itemize}
		\item Skills: 
		\begin{itemize}
		 	\item Technomat(Tinkerer) or Navigate Enclosed (Navigator)
		 	\item Pick two: Endure(Athlete), Jury Rig (Technomat), Calculus (Rationalist), Operate Void (Operator), Operate Aircraft (Operator), Break (Athlete), Cooking (Survivalist), Bookkeeping (Trader), Preach(Believer)
		 \end{itemize} 
	\end{itemize}

	% section lifepath_2_Homeworld (end)

\section{Step 3: Childhood} % (fold)
	\label{sec:lifepath_3_childhood}
	Tightly coupled with the influence of your homeworld, is the way of your upbringing. Are you noble born? Raised in a familiar manner? Or rather grew up in a tightly knit community? All alone? This will have serious influence on your character.\\
	In this step you will gain your second Aptitude, a singular Attribute adjustment and a single Profession with two free skill points to spend in any associated skills. Available upbringings are:
	\begin{itemize}
		\item Noble
		\item Familiar
		\item Communal
		\item Strict
		\item Neglected
		\item Loner
	\end{itemize}

	\subsection{Noble}
	\textit{You were born into a noble family. Etiquette, controlled demeanor and careful scheming was expected of you from a young age. Whether you lived up to that expectations or not is your call.}
	\begin{itemize}
		\item Attribute: Finesse+3 or Charisma+3
		\item Aptitude: Finesse or Charisma
		\item Profession: Trader, Believer, Betrayer, Commander, Spokesman, Rationalist
	\end{itemize}

	\subsection{Familiar}
	\textit{You were raised in a more or less loving and intimate family of the middle or even low class of your homeworld. Given just the right amount of love, guidance and freedom, you have grown into a likeable and free-willed person. Lucky you!}
	\begin{itemize}
		\item Attribute: Charisma+3 or Will+3
		\item Aptitude: Charisma or Will
		\item Profession: any Social or Intellectual (excluding Technologist)
	\end{itemize}

	\subsection{Communal}
	\textit{You were primarily raised by a close-knit, local community. Whether you are an orphan or your parents were working most of every day, the fact remains that you spent most of your early years with members of your local community: be they beggars, elders, gangers, artisans or merchants.}
	\begin{itemize}
		\item Attribute: Will+3 or Finesse+3
		\item Aptitude: Will or Finesse
		\item Profession: any Social, Warrior, Athlete, Acrobat, Survivalist
	\end{itemize}

	\subsection{Strict}
	\textit{You had a harsh childhood: strict parents, beating teachers, rough neighborhood, harsh rules and harsher punishments. Only by hard work did you ever managed to get the tiniest bit of recognition. You are like highly focused to reaching goals defined by your superiors no matter the cost and equally ready to punish you subordinates. Or you really have grown to hate just that kind of treatment.}
	\begin{itemize}
		\item Attribute: Toughness+3 or Smarts+3
		\item Aptitude: Toughness or Smarts
		\item Profession: Enforcer, Betrayer, any Physical or Intellectual (excluding Technologist)
	\end{itemize}

	\subsection{Neglected}
	\textit{You grew up neglected by your parents, family and the majority of the local community. You had to get by mostly alone, trusting in brute strength and instinct above all. At best you (eventually) decided that you'd do your damned best to rise above your origin. At worst, you took to the bad habits and behavior of those around you.}
	\begin{itemize}
		\item Attribute: Power+3 or Instinct+3
		\item Aptitude: Power or Instinct
		\item Profession: Survivalist, Enforcer, Betrayer, any Physical
	\end{itemize}

	\subsection{Loner}
	\textit{Loneliness my old friend. You grew up with only yourself to rely upon. Whether you were surrounded by indifferent strangers or literally alone in the wild does matter little. You had to learn quickly to listen to your instincts and senses when you wanted to life.}
	\begin{itemize}
		\item Attribute: Instinct+3 or Power+3
		\item Aptitude: Instinct or Power
		\item Profession: any Instinctual, any Physical
	\end{itemize}

	% section step_3_childhood (end)

\section{Step 4: Education} % (fold)
	\label{sec:lifepath_4_education}
	Ones education (in a general sense, not bound to classical classes!) leaves a lasting mark and enables a character to tackle certain areas with greater ease than others.\\
	This steps provides the Profession Group Aptitude, another two Attribute adjustments (one positive, the other negative) and two Professions - each with two free skill points to be spent within them! Available Educations:
	\begin{itemize}
		\item Schola Progenium
		\item Classic Education
		\item Mentor
		\item Self-Taught
		\item Indoctrinated
	\end{itemize}

	\subsection{Schola Progenium}
	\textit{The Eclesiarchy run Schola Progenium is a galaxy spanning institution to take in the orphans of high ranking Imperial citizens, especially of noble origin or those in the officer cadre of either Imperial Guard or Navy. The orphans receive one of the harshest physical and mental drilling and shaping known in the Imperium to forge them into perfect material for the next generation of stormtroopers, Adeptus Sororitas, Commissariat or Interrogators destined to join the Inquisition.}
	\begin{itemize}
		\item Group Aptitude: Physical or Social
		\item Attributes:
		\begin{itemize}
			\item Toughness+3 or Power+3
			\item Smarts-3 or Charisma-3
		\end{itemize}
		\item Professions: (pick two) Athlete, Believer, Commander, Enforcer, Warrior, Loremaster

	\end{itemize}

	\subsection{Classical Education}
	\textit{You received a year-long and well laid out schooling by (mostly able and sometimes nice) teachers together with many others of your age. The subjects were many and diverse, ranging from mathematics to accounting, biology, chemistry and philosophy. In fact, your young brain was stuffed beyond its capacity and you only recall some of it.}
	\begin{itemize}
		\item Group Aptitude: Intellectual
		\item Attributes:
		\begin{itemize}
			\item Smarts+3
			\item Power-3 or Finesse-3
		\end{itemize}
		\item Professions:
		\begin{itemize}
			\item Loremaster
			\item Spokesman, Navigator, Rationalist, Trader
		\end{itemize}
	\end{itemize}

	\subsection{Mentored}
	\textit{You received your education by a personal mentor in intimate sessions, learning from his experience and vast knowledge. Depending on your origin, this Mentor might be a specialized palace Tutor, payed highly by your noble family to teach you the rules of nobility or it might be the elderly shaman teaching the young of your tribal clan the ways of the word and the laws of nature.}
	\begin{itemize}
		\item Group Aptitude: Intellectual or Social
		\item Attributes:
		\begin{itemize}
			\item Any+3
			\item Any-3
		\end{itemize}
		\item Professions:
		\begin{itemize}
			\item Loremaster or Spokesman or Survivalist
			\item any Intellectual, any Social, any Instinctual
		\end{itemize}
	\end{itemize}

	\subsection{Self-Taught}
	\textit{You did not have the luxury of any kind of dedicated schooling. Instead you had to learn all the hard lessons of live by yourself. But learning them you did, leaving you (likely) scarred but well able to know your limits and how to accustom to new environments.}
	\begin{itemize}
		\item Group Aptitude: Physical or Instinctual
		\item Attributes:
		\begin{itemize}
			\item Instinct+3
			\item Finesse-3 or Smarts-3
		\end{itemize}
		\item Professions:
		\begin{itemize}
			\item Survivalist, Betrayer, Tinkerer
			\item any Instinctual, any Physical
		\end{itemize}
	\end{itemize}

	\subsection{Indoctrinated}
	\textit{You received extensive psychic indoctrination by the ill-understood tutolage-engines of the dark age of technology or other (potentially more sinister) means. Whatever the means, your mind is filled with knowledge and know-how of things you never truly learned or experienced. You just know them. Sometimes, things just comes to you as you act, without knowing you can actually do or know it beforehand.}
	\begin{itemize}
		\item Group Aptitude: Intellectual or Instinctual
		\item Attributes:
		\begin{itemize}
			\item Smarts+3 or Will+3 or Finesse+3
			\item Finesse-3 or Will-3
		\end{itemize}
		\item Professions: (pick two) any Intellectual, any Instinctual
	\end{itemize}

	% section step_4_education (end)


\section{Step 5: Combat} % (fold)
	\label{sec:lifepath_5_combat}
	In a future where there is only war, your character is nearly guaranteed to have encountered fighting before - either real or only simulated.\\
	This step will provide three skill points.

	\subsection{Gang War}
	You participated in a prolonged, bloody gang war, either as a gang member, militia or local law enforcement. Fighting dirty was the norm and no man or woman a non-combatant.
	\begin{itemize}
		\item Walking Fire (Gunman) or Hammerblow (Warrior)
		\item Pick two: Bribe (Trader), Endure (Athlete), Warcry (Enforcer), Hide (Ghost), Demolitions (Tinkerer), Wheeled (Operator), Gunner (Operator), Danger Sense (Observer), any Betrayer
	\end{itemize}

	\subsection{Hostile Wildlife}
	You've been to a place with lethal predators. You have learned your lessons from the grisly demise of others who didn't.
	\begin{itemize}
		\item Sniping (Gunman) or Piercing Strike (Warrior) or Brawl (Athlete)
		\item Pick two: First Aid (Medic), Diagnose (Medic), Silent Move (Ghost), Hide (Ghost), Sixth Sense (Observer), any Survivalist 
	\end{itemize}

	\subsection{PDF Training}
	You, as countless billion other imperial citizens, have taken the PDF training. Whether as for active service or only as part of auxiliary training programs is irrelevant, as the basic training regime is the same. Physical toughening through harsh exercise, weapon training, basic close quarter drills, life ammunition combat scenarios.
	\begin{itemize}
		\item Center Mass (Gunman) or Walking Fire (Gunman)
		\item Fencing(Warrior), Brawl (Athlete), Athletics (Athlete), Throw (Athlete), Hard Target (Acrobat)
		\item Technomat(Tinkerer), Demolition(Tinkerer), Bookkeeping (Trader), First Aid (Medic), Rally (Commander), Pull Rank (Commander), any Operator
	\end{itemize}

	\subsection{Showfighter}
	Whether you were a famous gladiator on a feudal world, an enslaved prisoner of war forced to fight to entertain your capturers, a boxing champion, or a fighter in an underhive pit: you made a living off fighting flashy, potentially highly choreographed battles for the entertainment of others.
	\begin{itemize}
		\item Fencing (Warrior), Brawl (Athlete), Combat Trick (Warrior) skill
		\item Hard Target (Acrobat), Jump (Acrobat), Endure (Athlete), Block (Warrior)
		\item Feint (Betrayer), Maneuver (Acrobat), Break (Athlete), Athletics (Athlete)
	\end{itemize}

	\subsection{Warzone}
	War has come to your home. Maybe your homeworld was invaded by foul Xenos or a rebellion has broken out. Maybe your void ship was boarded. Either way: you had to learn fast that staying alive is the most urgent thing to do when combat starts.
	\begin{itemize}
		\item Snap Shot (Gunman), Fencing (Warrior)
		\item Block(Warrior), Endure (Athlete) or Hard Target (Acrobat)
		\item Any Observer, any Survivalist
	\end{itemize}

	\subsection{None}
	You are a truly fortunate soul to have never experienced combat (not even simulated) in your life - a true rarity in the Grimdark future! But woe to you when combat eventually finds you...
	\begin{itemize}
		\item Any two Loremaster skills
		\item Free skill point (combat unrelated)
	\end{itemize}
	% section step_5_combat (end)

\section{Step 6: Early Career} % (fold)
	\label{sec:lifepath_6_early_career}
	Your first "job" left you with a deeper set of skills in a single field of expertise and is a good stepping stone to build upon in the future. Note that your early career does not necessarily reflect your character's current career.\\
	This stage provides another two stat adjustments (both positive) and a single Profession with three skill points.
	Available options:
	\begin{itemize}
		\item Adeptus Mechanicus
		\item Adeptus Ministorum (Ecclesiarchy)
		\item Adeptus Arbites
		\item Adeptus Administratum
		\item Soldier
		\item Enlisted Officer
		\item Criminal
		\item Outcast
		\item Civilian
	\end{itemize}

	\subsection{Adeptus Mechanicus}
	\textit{You have served as an adept in the marsian cult for several years. Countless hours spent surrounded by whirring, clicking, steaming, pumping and crackling machines, fixing corroded cabling, rusted pipes or replacing parts of degrading servitors while studying the secrets of the machine God in preparation for your next exams to rise up its many, complex ranks. }
	\begin{itemize}
		\item Stats:
		\begin{itemize}
			\item Smarts+3
			\item Will+3 or Finesse+3
		\end{itemize}
		\item Profession: Technologist
	\end{itemize}

	\subsection{Adeptus Ministorum}
	\textit{You've been sanctioned part of the most holy Ecclesiarchy, trained and educated to be a fine (or threatening) example for the masses to aspire to (or fear). Preaching the divine Emperor's name and bearing his golden lights to the deepest shadows has been your duty for years, leaving you either highly fervent or at the very least very charismatic and accustomed to talking to people of any calling.}
	\begin{itemize}
		\item Stats:
		\begin{itemize}
			\item Charisma+3
			\item Will+3 or Power+3
		\end{itemize}
		\item Profession: Believer or Spokesman
	\end{itemize}

	\subsection{Adeptus Arbites}
	\textit{You have been taken in by the Adeptus Arbites, the imperial law enforcement, tasked to uphold his majesty's law and order. Putting down rioters was as much part of your duty as fighting organized crime and misbehaving officials. It is generally a hard, ill-regarded duty with the potential to make numerous deadly foes quickly.}
	\begin{itemize}
		\item Stats:
		\begin{itemize}
			\item Instincts+3
			\item Any+3
		\end{itemize}
		\item Profession: Athlete, Enforcer, Observer
	\end{itemize}

	\subsection{Adeptus Administratum}
	\textit{You have learned and worked in the single largest bureaucracy ever to exist, keeping the Imperium in its quasi-functional status quo. With a single stroke of an autoquill you send armies on their way, condemned billions to their deaths and signed treaties and laws you were certain would be broken before your next shift.}
	\begin{itemize}
		\item Stats:
		\begin{itemize}
			\item Smarts+3 or Instinct+3
			\item Any+3
		\end{itemize}
		\item Profession: Trader, Loremaster, Rationalist
	\end{itemize}

	\subsection{Soldier}
	\textit{You have been drafted or chosen to join the massive fighting force of humanity, fighting the never ending wars of the Imperium. A harsh life full of hardship, pain and terror but also camaraderie, honor and wonder. }
	\begin{itemize}
		\item Stats:
		\begin{itemize}
			\item Toughness+3
			\item Power+3 or Finesse+3
		\end{itemize}
		\item Profession: Warrior, Gunman, Operator
	\end{itemize}

	\subsection{Enlisted Officer}
	\textit{You have gained (a low to medium) officer rank in one the Imperiums fighting forces. Whether you were elevated from the enlisted troops or went through one of many renowned officer schools is your call.}
	\begin{itemize}
		\item Stats:
		\begin{itemize}
			\item Smarts+3 or Charisma+3
			\item Will+3 or Instinct+3
		\end{itemize}
		\item Profession: Commander or Enforcer
	\end{itemize}

	\subsection{Criminal}
	\textit{In many places of the Imperium, illicit actions proof to be highly lucrative. From kidnapping and ransoming highborns to trafficking of illegal goods to industrial or social espionage to hired hitmans in service to some wealthy elements seeking to get rid of their contestants.}
	\begin{itemize}
		\item Stats:
		\begin{itemize}
			\item Finesse+3 or Charisma+3
			\item +3 to any except Will
		\end{itemize}
		\item Profession: Betrayer, Ghost
	\end{itemize}

	\subsection{Outcast}
	\textit{Living on or beyond the edge of civilization for several years, you have learned to only depend on yourself. You are highly praised as guide and expert by those having to survive in just such an area with little to no resources - although you will likely stick out like a colorful canine on any social gathering. }
	\begin{itemize}
		\item Stats:
		\begin{itemize}
			\item Instinct+3
			\item Toughness+3 or Power+3
		\end{itemize}
		\item Profession: Survivalist
	\end{itemize}

	\subsection{Civilian}
	\textit{You have learned and worked an utterly normal, civilian job for some years. }
	\begin{itemize}
		\item Stats:
		\begin{itemize}
			\item Finesse+3
			\item Any+3
		\end{itemize}
		\item Profession: Trader, Spokesman, Athlete, Acrobat, Operator, Medic, Navigator, Tinkerer, Loremaster
		\item Note: make sure to note down which was you civilian job and check in with your GM about which Profession is fitting. Examples:
		\begin{itemize}
			\item Medic: Nurse, Doctor, Veterinary
			\item Loremaster: Librarian, Scribe, Secretary
			\item Spokesman: Butler, Courtesan, Storyteller
			\item Operator: trucker, shuttle pilot, cargo train worker
		\end{itemize}
	\end{itemize}
% section step_6_early_career (end)